---
title: "Exploring Palmer Penguins using R"
author: "Jiyuan Wang"
date: "2026-09-22"
description: "An exploration of penguin body mass across species using R."
---

## Data

Data: [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/), Palmer Station Antarctica LTER.

## Load the data

```{r}
#| echo: true
#| message: false
#| warning: false

library(tidyverse)
library(palmerpenguins)

penguins |> head()
```

## Average body mass by species

```{r}
#| echo: true
penguins |>
  drop_na(species, body_mass_g) |>
  group_by(species) |>
  summarise(
    mean_body_mass = mean(body_mass_g),
    n = n()
  )
```

Gentoo penguins have the largest average body mass among the three species in this dataset.

## Body mass distribution by species

```{r}
#| echo: true
#| label: fig-body-mass-species
#| fig-cap: "Distribution of body mass for each penguin species."

penguins |>
  drop_na(species, body_mass_g) |>
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot() +
  labs(
    x = "Species",
    y = "Body mass (g)",
    title = "Penguin Body Mass by Species"
  )
```

The boxplot shows that Gentoo penguins generally have a higher body mass than Adelie and Chinstrap penguins.

## Flipper length and body mass

Next, I explored whether penguins with longer flippers also tend to have greater body mass.

```{r}
#| echo: true
#| label: fig-flipper-mass
#| fig-cap: "Relationship between flipper length and body mass by species."

penguins |>
  drop_na(flipper_length_mm, body_mass_g, species) |>
  ggplot(
    aes(
      x = flipper_length_mm,
      y = body_mass_g,
      color = species
    )
  ) +
  geom_point(alpha = 0.7) +
  labs(
    x = "Flipper length (mm)",
    y = "Body mass (g)",
    color = "Species",
    title = "Flipper Length vs Body Mass"
  )
```

The scatter plot shows a positive relationship between flipper length and body mass. Gentoo penguins tend to have both longer flippers and higher body mass than Adelie and Chinstrap penguins.


## Average flipper length by species

I also calculated the average flipper length for each species.

```{r}
#| echo: true

penguins |>
  drop_na(species, flipper_length_mm) |>
  group_by(species) |>
  summarise(
    mean_flipper_length = mean(flipper_length_mm),
    median_flipper_length = median(flipper_length_mm),
    n = n()
  )
```

Gentoo penguins have the largest average flipper length, while Adelie penguins generally have the shortest flippers.


## Compare average flipper length

The differences in average flipper length can also be visualized with a bar chart.

```{r}
#| echo: true
#| label: fig-average-flipper
#| fig-cap: "Average flipper length of each penguin species."

penguins |>
  drop_na(species, flipper_length_mm) |>
  group_by(species) |>
  summarise(
    mean_flipper_length = mean(flipper_length_mm)
  ) |>
  ggplot(
    aes(
      x = species,
      y = mean_flipper_length,
      fill = species
    )
  ) +
  geom_col() +
  labs(
    x = "Species",
    y = "Average flipper length (mm)",
    title = "Average Flipper Length by Species"
  ) +
  guides(fill = "none")
```

The bar chart confirms that Gentoo penguins have noticeably longer flippers on average than the other two species.